\scshape Topics in mechanism design

### 2025-02-04 ch1.tex
Add files via upload

% Copyright (c) 2024 Carl Martin Ludvig Sinander.

@@

@@

Here's a very simple (indirect) mechanism that the monopolist could adopt: post a price!
More fully, the monopolist sets a price $\pi \in \R_+$
and gives the agent two options:
purchase the good at price $\pi$ (pay $\pi$ and get the good for sure)
or don't (pay nothing and get nothing).

@@

\begin{exercise}
	%
	\label{exercise:posted_price}
	%
	(a) This allocation--payment pair (viewed as a direct mechanism) is IC; why?
	(b) Verify that this mechanism satisfies the envelope formula.
	(c) There are other mechanisms $(q,p')$, i.e. mechanisms with the same allocation but different payments. Describe them.
	%
\end{exercise}

>>>>

% Copyright (c) 2025 Carl Martin Ludvig Sinander.

@@

\begin{exercise}
	%
	\label{exercise:first_price}
	%
	Consider the (single-agent) first-price auction with reserve price $\pi \in \R_+$: the agent submits a `bid' $b \in \R_+$, where in case $b \geq \pi$ she pays her bid $b$ and gets the good for sure, while if $b < \pi$ then she pays nothing and gets nothing. This is a mechanism.
	%
	\begin{enumerate}[label=(\alph*)]

		\item What are the allocation and payment rule $\left(q^1,p^1\right)$ induced by this mechanism?

		\item Is the DRM $\left(q^1,p^1\right)$ incentive-compatible? Why or why not?

\begin{exercise}
	%
	\label{exercise:posted_price}
	%
	Consider the `posted-price' mechanism whereby the principal sets a price $\pi \in \R_+$ and gives the agent two options: purchase the good at price $\pi$ (pay $\pi$ and get the good for sure) or don't (pay nothing and get nothing). This is a mechanism. It can be viewed as a single-bidder second-price auction with reserve price $\pi$.%
		\footnote{Let the agent submit a `bid' $b$; if $b \geq \pi$ she pays the reserve price $\pi$ and gets the good for sure, while if $b < \pi$ she pays nothing and gets nothing.}
	Alternatively, it can be viewed as a direct revelation mechanism, namely $(q,p)$ given by
	%
	\begin{equation*}
		(q(t),p(t)) \coloneqq
		\begin{cases}
			(0,0)	& \text{if $t<\pi$} \\
			(1,\pi)	& \text{if $t \geq \pi$.}
		\end{cases}
	\end{equation*}
	%
	\begin{enumerate}[label=(\alph*)]

		\item What are the allocation and payment rule $\left(q^2,p^2\right)$ induced by this mechanism?

		\item Is the DRM $\left(q^2,p^2\right)$ incentive-compatible? What about the DRM $(q,p)$?

		\item What is the relationship between the DRM $(q,p)$ and the DRM $\left(q^1,p^1\right)$ from \Cref{exercise:first_price} above (the one induced by the first-price auction with reserve price $\pi$)?

@@

Here's a very simple (indirect) mechanism that the monopolist could adopt: post a price!
More fully, the monopolist sets a price $\pi \in \R_+$
and gives the agent two options:
purchase the good at price $\pi$ (pay $\pi$ and get the good for sure)
or don't (pay nothing and get nothing). (We saw this mechanism in \Cref{exercise:posted_price} near the start of this chapter.)

@@

\begin{exercise}
	%
	\label{exercise:posted_price2}
	%
	(a) This allocation--payment pair (viewed as a direct mechanism) is IC; why?
	(b) Verify that this mechanism satisfies the envelope formula.
	(c) There are other mechanisms $(q,p')$, i.e. mechanisms with the same allocation but different payments. Describe them.
	%
\end{exercise}

### 2025-02-04 ch2.tex
Add files via upload

\paragraph{The literature.}
The symmetric realisability problem was studied first.
\textcite{MaskinRiley1984} took the first steps.
\textcite{Matthews1984} proved the necessity of the Border inequalities, and conjectured sufficiency.
The conjecture was proved by \textcite{Border1991}, using the above argument above based on hierarchical allocations and a separating hyperplane.
\textcite{Border2007} obtained the asymmetric version for finitely many types,
and \textcite{Mierendorff2011} extended this to general type spaces.
\textcite{HartReny2015border} provide an alternative characterisation of realisability in terms of second-order stochastic dominance.
Some (not many) extensions to richer environments have been considered; for example, \textcite{CheKimMierendorff2013} allow for certain sorts of constraints.

@@

This is not a linear programme, because $Q_i$ also enters non-linearly.
But we can convert it into a parametrised linear programme:
given weights $\varphi_1,\dots,\varphi_I \in [0,1]$ that sum to $\leq 1$, consider maximising
%
\begin{equation*}
	\sum_{i=1}^I \E \left[ Q_i(T_i) [T_i-c_i]
	+ c_i \varphi_i \right] 
\end{equation*}
%
subject $Q_i \geq \varphi_i$ and the Border inequalities.
It can be shown by an elementary argument \parencite[see][]{ErlansonKleiner2019} that whatever the values of $(\varphi_i)_{i=1}^I$, any solution of this linear programme is a \emph{threshold mechanism.}
Thus optimal mechanisms must be threshold mechanisms.

>>>>

\paragraph{The literature.}
The symmetric realisability problem was studied first.
\textcite{MaskinRiley1984} took the first steps.
\textcite{Matthews1984} proved the necessity of the Border inequalities, and conjectured sufficiency.
The conjecture was proved by \textcite{Border1991}, using the above argument based on hierarchical allocations and a separating hyperplane.
\textcite{Border2007} obtained the asymmetric version for finitely many types,
and \textcite{Mierendorff2011} extended this to general type spaces.
\textcite{HartReny2015border} provide an alternative characterisation of realisability in terms of second-order stochastic dominance.
Some (not many) extensions to richer environments have been considered; for example, \textcite{CheKimMierendorff2013} allow for certain sorts of constraints.

@@

This is not a linear programme, because $Q_i$ also enters non-linearly.
But we can convert it into a parametrised linear programme:
given weights $\varphi_1,\dots,\varphi_I \in [0,1]$ that sum to $\leq 1$, consider maximising
%
\begin{equation*}
	\sum_{i=1}^I \E \left[ Q_i(T_i) [T_i-c_i]
	+ c_i \varphi_i \right] 
\end{equation*}
%
subject to $Q_i \geq \varphi_i$ and the Border inequalities.
It can be shown by an elementary argument \parencite[see][]{ErlansonKleiner2019} that whatever the values of $(\varphi_i)_{i=1}^I$, any solution of this linear programme is a \emph{threshold mechanism.}
Thus optimal mechanisms must be threshold mechanisms.

### 2025-02-04 ch3.tex
Add files via upload

\paragraph{The literature.} \textcite{LipnowskiRamos2020} characterise principal-optimal equilibria (and more generally, Pareto-optimal equilibria).%
	\footnote{In their model, the agent learns $T_n$ only if and when she is offered the good in period $n$. This does not make a difference, I believe.
	For suppose the principal-preferred equilibrium were to feature the agent conditioning her behaviour in period $n+1$
	on what $T_n$ if she was not offered the good in period $n$.
	Since $T_n$ has no statistical effect on $T_{n+1}$ (that's our IID assumption),
	this just amounts to the agent randomising in period $n+1$ (when the principal did not offer the good in period $n$);
	so we can replicate this behaviour by having the agent mix (i.e. she can use her own set of dice to randomise, instead of conditioning her behaviour on the `natural dice' $T_n$).}
Their analysis relies on the `APS' technique \parencite{AbreuPearceStacchetti1990}, which provides a recursive formulation of repeated games with imperfect monitoring. This is a standard (indeed, essential!) tool for analysing repeated interactions without commitment.
The original APS paper is rather beautiful,
but you may prefer to learn it from the textbook treatment of \textcite{MailathSamuelson2006}, or from whatever lecture notes you can find around the internet.
Another nice paper on dynamic allocation without commitment is \textcite{DeclippelEliazFershtmanRozen2021}, who study a model with \emph{two} agents.

>>>>

\paragraph{The literature.} \textcite{LipnowskiRamos2020} characterise principal-optimal equilibria (and more generally, Pareto-optimal equilibria).%
	\footnote{In their model, the agent learns $T_n$ only if and when she is offered the good in period $n$. This does not make a difference, I believe.
	For suppose the principal-preferred equilibrium were to feature the agent conditioning her behaviour in period $n+1$
	on what $T_n$ was if she was not offered the good in period $n$.
	Since $T_n$ has no statistical effect on $T_{n+1}$ (that's our IID assumption),
	this just amounts to the agent randomising in period $n+1$ (when the principal did not offer the good in period $n$);
	so we can replicate this behaviour by having the agent mix (i.e. she can use her own set of dice to randomise, instead of conditioning her behaviour on the `natural dice' $T_n$).}
Their analysis relies on the `APS' technique \parencite{AbreuPearceStacchetti1990}, which provides a recursive formulation of repeated games with imperfect monitoring. This is a standard (indeed, essential!) tool for analysing repeated interactions without commitment.
The original APS paper is rather beautiful,
but you may prefer to learn it from the textbook treatment of \textcite{MailathSamuelson2006}, or from whatever lecture notes you can find around the internet.
Another nice paper on dynamic allocation without commitment is \textcite{DeclippelEliazFershtmanRozen2021}, who study a model with \emph{two} agents.

### 2025-02-04 ms.tex
Add files via upload

\date{\small This version: 8 November 2024}

@@

\pagebreak
\hspace{1pt}\vfill
\noindent
Copyright \copyright{} 2024 Carl Martin Ludvig Sinander.

@@

These are the notes for a second-year graduate course that I have been teaching at Oxford each autumn since 2021.
Thanks to Gregorio Curello, Péter Es\H{o}, Deniz Kattwinkel, Elliot Lipnowski, Qingmin Liu, Alex Teytelboym and to my students for comments, suggestions and insights,
and to Ufuk Altunbüken, Jo Auer, Catharina Behrens, Josh le Cornu, Tanxin Gao, Justin Hadad, Brooklyn Han, Simon Handreke, Sanjari Kalantri, Marie Kaul, Nils Lager, Arvo Munoz Moran, Sami Petersen, Filip Tokarski, Gautam Vyas and Yidan Xu for reporting errors.

@@

\setcounter{example}{1}
\begin{example}
	%
	\label{example:fg}
	%
	Consider the vector space of functions $[0,1] \to \R$.
	Let $C$ be the set of all functions $h : [0,1] \to \R$ which have the form
	$f(x) = k + (2-k) x$
	for some $k \in [0,2]$.
	(Try drawing a picture to visualise $C$.)
	This is a convex set. (Prove it!)
	%
\end{example}

>>>>

\date{\small This version: 4 February 2025}

@@

\pagebreak
\hspace{1pt}\vfill
\noindent
Copyright \copyright{} 2025 Carl Martin Ludvig Sinander.

@@

These are the notes for a second-year graduate course that I have been teaching at Oxford each autumn since 2021.
Thanks to Gregorio Curello, Péter Es\H{o}, Deniz Kattwinkel, Elliot Lipnowski, Qingmin Liu, Alex Teytelboym and to my students for comments, suggestions and insights,
and to Ufuk Altunbüken, Jo Auer, Catharina Behrens, Josh le Cornu, Tanxin Gao, Justin Hadad, Brooklyn Han, Simon Handreke, Sanjari Kalantri, Marie Kaul, Nils Lager, Arvo Munoz Moran, Sami Petersen, Augustus Smith, Filip Tokarski, Gautam Vyas and Yidan Xu for reporting errors.

@@

\setcounter{example}{1}
\begin{example}
	%
	\label{example:fg}
	%
	Consider the vector space of functions $[0,1] \to \R$.
	Let $C$ be the set of all functions $f : [0,1] \to \R$ which have the form
	$f(x) = k + (2-k) x$
	for some $k \in [0,2]$.
	(Try drawing a picture to visualise $C$.)
	This is a convex set. (Prove it!)
	%
\end{example}
